Structural replication of nanoporous media using procedural noise

### 2017-06-07 ms.tex
Har lagt til referanser og gjort noen smaa endringer i introduction. Det som naa mangler er en litt mer utfyllende diskusjon.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract} 
In this paper, we aim to develop a method for producing statistically correct porous media using procedural noise functions. In short, the proposed model is a binary function that returns either true if the input point is inside pore space, or false if within a wall. By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales. It is therefore possible to generate fractal-like geometries at arbitrary scales, circumventing the need for any physics simulations. In addition, this method is extremely fast and has low memory footprints, requiring very little computational power. After presenting such a model, we choose a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g. molecular dynamics) or with our method. The radial distribution function $g(r)$ is chosen as a measure of the structural properties to compare the simulated geometry to our model. Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters. Finally, we generate a nanoporous media with an expand-and-quench technique using molecular dynamics and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area. 
  \todo{Abstract boer tenkes noe gjennom: faa frem hovedpoenger.}

\end{abstract}

@@

The rapid development of new experimental and observational methods to determine the three-dimensional structure of many types of porous media calls for new statistical tools and methods to characterize complex porous media and to generate synthetic media based on the observed structures.
New experimental methods allow for studies of three-dimensional structures from nano-meter to meter scales.
Destructive techniques such as focused ion beam scanning electron microscopy (FIB-SEM) are routinely used to characterize geologically relevant structures in shales and carbonate rocks with resolutions down to tens of nanometers \cite{XX}.
High-energy X-ray tomography methods using high-energy beams at international research facilites or in-house scanners open for three-dimensional mapping of mineral structures from sub-micrometer to millimeter scales \cite{XX}.
Thus new experimental techniques provide three-dimensional voxelized images that can form the basis of analysis or modeling.
However, in many cases we can only scan a few examples experimentally, but need to perform simulations over many statistically similar realizations to find effective properties or to populate upscaling models.
Thus the need for methods to robustly characterize three-dimensional porous structures and to generate synthetic structures with similar statistical properties.

The field of synthetic rocks is a large field both scientifically and technologically \cite{}. 
Methods to generate synthentic rock samples from experimental characterizations may be based on purely statistical or geometrical models, models of physical processes that generate the samples, or combinations of both. 
For small-scale porous media, physical models are often used to generate initial grain packings \cite{}, then further mechano-chemical and fluid-mediated processes such as precipitation-dissoluton processes and pressure solution \cite{}, or fracture and plastic deformation processes are added to include effects of long-term compaction \cite{}. 
On large scales, reservoir models are populated by statistical models for lithological and geomorphological processes \cite{}, such as river fan formation to generate particular grain size distributions \cite{}. 
More statistical approaches include variations of percolation models \cite{sahimi} or network models \cite{} to characterize and generate uncorrelated or correlated porous media. 
In addition, the high resolution of observational methods allow three-dimensional scans to form the basis for atomic-scale modeling using molecular modeling methods such as molecular dynamics or Monte Carlo methods to study the dynamics of nanoporous media. 
But then procedural or physical models are needed to fill in the atomic scale details, to build many statistical realizations of nanoporous media with atomic-scale details, and to extrapolate to larger scale models.

@@

In this paper, we have proposed a novel method for generating a nanoporous media using procedural methods in the form of simplex noise. 
Procedural methods, albeit having no root in physics, have some advatages over prevalent methods: they are computationally efficient, have low memory footprints and are only limited in terms of detail by the model in question. 
In particular, it is possible to create simulations of arbitrary scales and resolution, indeed with any level of detail.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract} 
New experimental and observational methods of the structure of three-dimensional porous media calls for improved approaches for statistical characterization and the generation of random samples with given statistical properties. 
We aim to develop a method for producing statistically consistent realization of porous media using procedural noise functions. By using procedural noise, we are not restricted to generating media with a predetermined resolution, since the noise functions is continuous on all scales. It is therefore possible to generate fractal-like geometries at arbitrary scales. This method is extremely fast and has low memory footprints, requiring very little computational power. We demonstrate the use of procedural noise functions by choosing a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g. molecular dynamics) or with our method. The radial distribution function $g(r)$ is chosen as a measure of the structural properties to compare the simulated geometry to our model. Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters. Finally, we generate a nanoporous media with an expand-and-quench technique using molecular dynamics and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area. 
\end{abstract}

@@

The rapid development of new experimental and observational methods to determine the three-dimensional structure of many types of porous media calls for new statistical tools and methods to characterize complex porous media and to generate synthetic media based on the observed structures.
New experimental methods allow for studies of three-dimensional structures from nano-meter to meter scales.
Destructive techniques such as focused ion beam scanning electron microscopy (FIB-SEM) are routinely used to characterize geologically relevant structures in shales and carbonate rocks with resolutions down to tens of nanometers \cite{cnudde2013high,curtis2010structural}.
High-energy X-ray tomography methods using high-energy beams at international research facilites or in-house scanners open for three-dimensional mapping of mineral structures from sub-micrometer to millimeter scales \cite{kobchenko20114d,jamtveit2014pore,zhu2011microtomography,fusseis2009creep}.
Thus new experimental techniques provide three-dimensional voxelized images that can form the basis of analysis or modeling.
However, in many cases we can only scan a few examples experimentally, but need to perform simulations over many statistically similar realizations to find effective properties or to populate upscaling models.
Thus the need for methods to robustly characterize three-dimensional porous structures and to generate synthetic structures with similar statistical properties.

The field of synthetic rocks is a large field both scientifically and technologically \cite{biswal2007stochastic,hamzehpour2006development,Reconstructing random media}. 
Methods to generate synthentic rock samples from experimental characterizations may be based on purely statistical or geometrical models, models of physical processes that generate the samples, or combinations of both. 
For small-scale porous media, physical models are often used to generate initial grain packings \cite{pilotti1998generation,scott1969density,herrmann2013physics}, then further mechano-chemical and fluid-mediated processes such as precipitation-dissoluton processes and pressure solution \cite{utter1983pressure}, or fracture and plastic deformation processes are added to include effects of long-term compaction \cite{rutter1983pressure,renard1997pressure}. 
On large scales, reservoir models are populated by statistical models for lithological and geomorphological processes \cite{pyrcz2014geostatistical}, such as river fan formation to generate particular grain size distributions \cite{oltermann1996heterogeneity,lair1994alluvial}. 
More statistical approaches include variations of percolation models \cite{sahimi} or network models \cite{fatt1956network,sahimi1993flow} to characterize and generate uncorrelated or correlated porous media. 
In addition, the high resolution of observational methods allow three-dimensional scans to form the basis for atomic-scale modeling using molecular modeling methods such as molecular dynamics or Monte Carlo methods to study the dynamics of nanoporous media. 
But then procedural or physical models are needed to fill in the atomic scale details, to build many statistical realizations of nanoporous media with atomic-scale details, and to extrapolate to larger scale models.

@@

In this paper, we have proposed a novel method for generating a nanoporous media using procedural methods in the form of simplex noise. 
Procedural methods are not physical, but rather statistical models, and this gives them some advatages over prevalent methods: they are computationally efficient, have low memory footprints and are only limited in terms of detail by the model in question. 
In particular, it is possible to create simulations of arbitrary scales and resolution, indeed with any level of detail.
